---
title: "Using PTRMSR for automatizing pre-processing"
output:
  pdf_document: default
  toc: true
---


\tableofcontents

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction and settings
This vignette shows how to use PTRMSR package to read .txt files, calculate AUC and return statistical analyses. 

```{r, eval=FALSE}
if (!require("devtools", quietly = TRUE))
    install.packages("devtools")

library(devtools)
install_github("https://github.com/ChemoSens/PTRMSR")
``` 

Then, the PTRMSR package should be loaded.

```{r, eval=FALSE}
library(PTRMSR)
``` 

# Step 1: Observing the data for one or two files
## Outputs for one given file
It could be interesting to see the raw data of one file. In this purpose, the function ptrReport returns several plots which could help to interpretation. 
```{r,eval=FALSE}
wd="C:/INRA/Data/Donnees Cantin/Cantin-DTS-PTRviewer" 
setwd(wd)
listFiles=list.files(pattern="*.txt")
referenceBreath="m69.06906 (69.06906) (Conc)"
file=listFiles[3]
dataset=read.table(file=file,header=TRUE,sep="\t",check.names=FALSE)
report=ptrvReport(dataset,selecIons="evolving",
                  listIons=NULL,
                  referenceBreath=referenceBreath,
                  methodDetectStart="startPeakProportion",
                  noisePeriodIBT=c(0,30),noisePeriodSig=c(0,30),
                  noisePeriodDS=c(0,30),
                  proportionOfMax=0.3,halfWindowSize=4,forMinExpiDivideMaxIntBy = 7,forMaxInspiDivideMaxIntBy =)
names(report$gg)
report$gg$p_breath$p_cyclelimits
report$gg$p_breath$p_smoothbreath
```

Regarding the breathing cycle detection, these two outputs gives the results of cycle limits and the smooth data for breathing (useful to adjust the parameters halfWindowSize and maxPeaks)

```{r,eval=FALSE}
plot(report$gg$p_breath$p_cyclelimits)
plot(report$gg$p_breath$p_smoothbreath)
```

Regarding the ions distributions
```{r,eval=FALSE}
plot(report$gg$p_curves$p_raw)
plot(report$gg$p_curves$p_cycle)
```

Regarding the breathing ion distribution
```{r,eval=FALSE}
plot(report$gg$p_curves$p_breath_raw)
plot(report$gg$p_curves$p_breath_cycle)
```

Do not hesitate to use plotly for interactive graphs
```{r,eval=FALSE}
library(plotly)
ggplotly(report$gg$p_curves$p_breath_raw)
```


# Step 2: Reading and filling meta-data

## Loading metadata
All the data files (.txt from ptrViewer) and the metadata file (.csv) should be in a single repository (wd)
```{r, eval=FALSE}
wd="C:/INRA/Data/Donnees Cantin/Cantin-DTS-PTRviewer" 
setwd(wd)
listFiles=list.files(pattern="*.txt")[-1]
metaData2=read.table("metaData2.csv",sep=";",header=T)
head(metaData2[,-c(2:3)])
```

## Checks


## Analysis of PTR-Viewer files

### Selection of relevant ions only
This function returns the names of ions whose maximal intensity is 3 times higher than the maximal intensity during the noise period. It returns a list containing (i) a list containing for each file the ratio maximal intensity during the tasting/maximal intensity during the noise period, (ii) a vector (intersection) containing the ions which are significant in all files (iii) a vector (union) containing the ions which are significant in at least one file
```{r, eval=FALSE}
setwd(wd)

sigIons=ptrvListSignificantSNRIons(listFiles=listFiles[1:4],
               metaData=metaData2,noisePeriod=c(0,25))
ionSigUnique=sigIons$union
```



To obtain further help for this (or any) function, enter: 
```{r, eval=FALSE}
?ptrvListSignificantSNRIons
```


## AUC calculations
This function allows statistics to be calculated for each file after breathing correction
```{r,eval=FALSE,warning=FALSE,message=FALSE,include=FALSE}
setwd(wd)
res_auc=ptrvListIntensityByTime(listFiles=listFiles[1:4],
                                metaData=metaData2,ions=ionSigUnique)
```

The results can be saved into csv files
```{r,eval=FALSE}
res_auc$listRes
write.table(file="auc.csv",sep=";",res_auc$listRes,row.names=F)
```

The stat option can be used to select the statistic.
```{r,eval=FALSE}
res_tmax=ptrvListIntensityByTime(listFiles=listFiles,metaData=metaData2,ions=ionSigUnique,stat="tmax")
```

## Statistical analysis
In this part, the evaluations can be normalized (a total abundance of 1 for each evaluation) or logged. 
### PCA of observations
```{r,eval=FALSE}
respca=ptrvListPCA(ptrvList$totalIntensity,dataType="productMeans",log=FALSE)
# rajouter projetions individuelles plutot qu'ellipses
plotPCAgg(respca,type="ind")
plotPCAgg(respca,type="var",text=FALSE)
p=plotPCAgg(respca,type="var",text=FALSE)
p
library(plotly)
ggplotly(p)

p=plotPCAgg(respca,type="cor1",n=80)
plotPCAgg(respca,type="cor2")

respca=ptrvListPCA(ptrvList$totalIntensity,dataType="raw")
p=plotPCAgg(respca,type="ind",text=FALSE)
ggplotly(p)
plotPCAgg(respca,type="var",text=FALSE)

```

### Anova results
```{r,eval=FALSE,warning=FALSE}
resanova=ptrvListAnova(ptrvList$totalIntensity[ptrvList$totalIntensity[,"ion"]=="m101.09",],normalizeByEval=FALSE)
```

### Heatmaps
```{r,eval=FALSE}
ptrvListHeatmap(df=ptrvList$totalIntensity,formula=as.formula(product+subject+rep~ion),
                fun.aggregate="mean",clusterRows=T,clusterCols=T)
ptrvListHeatmap(df=ptrvList$totalIntensity,formula=as.formula(product+subject+rep~ion),
                fun.aggregate="mean",clusterRows=T,clusterCols=T)
ptrvListHeatmap(df=ptrvList$totalIntensity,formula=as.formula(product+subject+rep~ion),
                fun.aggregate="max",clusterRows=T,clusterCols=T)
ptrvListHeatmap(df=ptrvList$totalIntensity,formula=as.formula(product+subject~ion),
                fun.aggregate="max",clusterRows=T,clusterCols=T,showRownames=T)
ptrvListHeatmap(df=ptrvList$totalIntensity,formula=as.formula(product+subject~ion),
                fun.aggregate="mean",normalization="row",clusterRows=T,clusterCols=T,showRownames=T)
ptrvListHeatmap(df=ptrvList$totalIntensity,formula=as.formula(product+subject~ion),
                fun.aggregate="mean",normalization="row",clusterRows=T,clusterCols=T,showRownames=T,annotationRow="subject")

```



## Appendix
### Help for filling meta-data
```{r, eval=FALSE}
ptrvCreateMetaFile(wd,subject=c(9,12),product=NULL,replicate=NULL,sep=";")
metaData=read.csv("metaData.csv",sep=";",header=TRUE,dec=",")
metaData[,"resp"]=referenceBreath
``` 
###[OPTIONAL] Completing meta-data file with the tasting time
```{r, eval=FALSE}
# courbes + lissage
 t0=tn=suj=ordre_produit=product=repet=finTS=rep(NA,nrow(metaData))
  names(t0)=listFiles
  for(i in 1:nrow(metaData))
  {
    print("dataset")
    print(i)
    dataset=read.table(listFiles[i],sep="\t",dec=dec_vec[i],header=T)
    dataset[,"RelTime"]=as.numeric(dataset[,"RelTime"])
   # dataset[,ions]=apply(dataset[,ions],2,as.numeric)
    res_intensity=ptrvIntensityByTime(dataset,ions=ionSigAll
                                      ,referenceBreath=referenceBreath,
                                      correction = "cycle",
                                      removeNoise=TRUE,breathRatio =FALSE,timeBlank=c(0,25),
                                    halfWindowSize=12,maxPeaks=30,method="SuperSmoother",total=FALSE)
    t0[i]=ptrvDetectStart(res=res_intensity,starts=ionSigAll, method="startPeakProportion",proportionOfMax=0.1,graph=FALSE,
                          noisePeriod=c(0,25),startPeriod=c(25,100))$tx
    tn[i]=max(res_intensity$res[,"time"])
    suj[i]=substr(listFiles[i],9,12)
    ordre_produit[i]=substr(listFiles[i],14,14)
    tds_raw_i=tds_raw[tds_raw[,"SubjectCode"]==suj[i]&tds_raw[,"Rang.produit"]==ordre_produit[i],]
    product[i]=unique(tds_raw_i[,"ProductCode"])
    repet[i]=unique(tds_raw_i[,"Replicate"])
    finTS[i]=max(tds_raw_i[,"Time"])
  }
metaData[,"miseEnBouche_s"]=t0;metaData[,"finPTR_s"]=tn;
metaData[,"finTimeSens_s"]=finTS
metaData[,"product"]=product;metaData[,"subject"]=suj;metaData[,"rep"]=repet
write.table(metaData,file="metaData2.csv",sep=";",row.names=F,dec=".")
```
