---
title: "Using PTRMSR for automatizing pre-processing"
vignette: "%\\VignetteIndexEntry{Vignette Title} %\\VignetteEngine{knitr::rmarkdown}
  %\\VignetteEncoding{UTF-8}\n"
output:
  html_document
---


\tableofcontents

This vignette shows how to use PTRMSR package to read .txt files and calculate some statistics 

## Preliminary work

### Installing PTRMSR packages

The package can be uploaded with the following instructions. It requires several external packages that should also beinstalled. 

```{r eval=FALSE}
# remove.packages("PTRMSR)
# install.packages("devtools")
# install.packages("BiocManager")
# BiocManager::install("rhdf5") #<- update packages: no for faster
# BiocManager::install("MSnbase")#<- update packages: no for faster
library(rhdf5)
library(MSnbase)
library(devtools)
install_github("https://github.com/ChemoSens/PTRMSR")
```

### General principle
In this package, the functions beginning by ptrv use a single dataset as input. The functions beginning by ptrvList use several files and a metaData file containing information about the different files. 
The objective of this paper is to use files produced by ptrViewer (or any other preprocessing of PTR-MS device returning files with ion, intensity and time as columns) then return curves and relevant statistics on the evaluations. 
Several biases can be corrected:  (i) the breathing of the subject by calculating statistics by breathing cycle (use the parameter correction="cycle"),(ii) smoothing can be applied (instead of breathing cycle correction) (iii) blank period during the evaluation that can be removed from the intensities (use removeNoise = TRUE and noisePeriod=...). 
Some statistics are returned.

### Loading the library
```{r,eval=TRUE}
library(ggplot2)
library(gridExtra)
library(PTRMSR)
```

## Pre-processing only one file
### Loading a dataset
The first step is to load a dataset whose colnames are 'time', then the different ions. Each evaluation is a line of the dataset. 

ptrv is an example object where the time column is 'RelTime', with two supplementary columns to remove (AbsTime and Cycle).


### Correcting the breathing (or not)
ptrvIntensityByTime allows the breath to be corrected by calculating the breathing cycles than calculating a statistic per breathing cycle (funAggreagte='mean' for average or funAggregate="max" for maximum for example)
```{r, eval =TRUE}
data(ptrv)
resptrv=ptrvIntensityByTime(dataset=ptrv,timeCol="RelTime",colToRemove=c("AbsTime","Cycle"),correction="cycle",referenceBreath="m69.06989..isoprene...Conc.",timePeriod=NULL,
             ions=NULL,funAggregate="mean",                     minExpi=NULL,maxInspi=NULL,smoothMethodBreath="MovingAverage",minimalDuration=2,forMinExpiDivideMaxIntBy=4,forMaxInspiDivideMaxIntBy=5,halfWindowSize=5, method="MAD",SNR=0)
names(resptrv)
resptrv$gg$p_cyclelimits
head(resptrv$res)
```

The obtained results contains a dataframe in a long format (as used in ggplot) whose colnames are time, intensity, ion and duration.
Running the same function with the parameter correction="none"  returns the raw data in a similar format without any correction of the breath. 
```{r, eval =TRUE}
resptrv_none=ptrvIntensityByTime(dataset=ptrv,timeCol="RelTime",colToRemove=c("AbsTime","Cycle"),correction="none")
head(resptrv_none$res)
```



### Smoothing
The resulting data can be smoothed using ptrvSmooth function. It allows different methods of smoothing to be tested.
```{r}
smoothed_df=ptrvSmooth(resptrv$res,method="Loess",spar=0.2,sameTime=FALSE)
```

```{r}
ggplot(smoothed_df,aes(x=time,y=intensity,col=ion))+geom_line()+theme(legend.position="none")
```

### Correcting by the noise

The time period can be corrected using ptrvRemoveNoise function

```{r, eval =TRUE}
df_noiseCorrected=ptrvRemoveNoise(smoothed_df,timeBlank=c(0,30))
```

```{r}
p=ggplot(df_noiseCorrected,aes(x=time,y=intensity,col=ion))+geom_line()+theme(legend.position="none")
#ggplotly(p)
```

### Visualization of all corrections for one given ion

This part allows the different steps of pretreatemnt to be vizualized. 
```{r, eval =TRUE}
p1=ggplot(ptrv,aes(y=m45.03349...C2H4O.H....Conc.,x=RelTime))+geom_line()+ggtitle("Raw data")

resion=resptrv$res[resptrv$res[,"ion"]=="m45.03349...C2H4O.H....Conc.",]
resion_none=resptrv_none$res[resptrv_none$res[,"ion"]=="m45.03349...C2H4O.H....Conc.",]
p2=ggplot(resion,aes(x=time,y=intensity))+geom_line()+ggtitle("breath correction")


p3=ggplot(smoothed_df[smoothed_df[,"ion"]=="m45.03349...C2H4O.H....Conc.",],aes(x=time,y=intensity))+geom_line()+ggtitle("breath correction + smooth")
smoothed_raw_df=ptrvSmooth(resion_none,method="Loess",spar=0.2,sameTime=FALSE)
p4=ggplot(smoothed_raw_df,aes(x=time,y=intensity))+geom_line()+ggtitle("smoothed raw data")
p_noise=ggplot(df_noiseCorrected[df_noiseCorrected[,"ion"]=="m45.03349...C2H4O.H....Conc.",],aes(x=time,y=intensity))+geom_line()+ggtitle("noise correction")
grid.arrange(p1,p2,p3,p4,p_noise)
```


## Managing several PTR-Viewer files with noise and breathing correction

### Loading metadata and list of files to be processed
All the data files (.txt from ptrViewer) and the metadata file (.csv) should be in a single repository 
```{r, eval=FALSE}
repo="./../inst/extdata"
listFiles=list.files(repo,pattern="*.txt")
metaData=read.table(paste0(repo,"/metaData.csv"),sep=";",header=T)
head(metaData)
```



### AUC calculations of several PTR-Viewer files with noise and breathing correction
This function allows statistics to be calculated for each file after breathing correction
```{r,eval=FALSE}
setwd(repo)
res_auc=ptrvListIntensityByTime(listFiles=listFiles,
                                metaData=metaData,correction="cycle", ions=c("m31.0183", "m31.0259", "m33.0324"))
head(res_auc$listRes)
```

Parameters can be added using smoothing=TRUE then smoothing parameters and removeNoise=TRUE then removeNoise parameters (time)
The results can be saved into csv files
```{r,eval=FALSE}
write.table(file="auc.csv",sep=";",res_auc$listRes,row.names=F)
```

The stat option can be used to select the statistic.
```{r,eval=FALSE}
repo="./../inst/extdata"
res_tmax=ptrvListIntensityByTime(listFiles=listFiles,metaData=metaData,ions=c("m31.0183", "m31.0259",                                                "m33.0324"),correction="cycle",stat="tmax")
```

The breathing cycles can be observed with this command.
```{r,eval=FALSE}
grid.arrange(grobs=res_tmax$cycleLimits[1:4])
```

## Documentations on the functions are available in the R documentation
```{r,eval=FALSE}
?detectCycle
?ptrvIntensityByTime # for only one dataset, 
?ptrvIntensity # for summarizing the times of one dataset
?ptrvSmooth
?ptrvRemoveNoise
?ptrvListIntensityByTime # for only one dataset, 
?ptrvListIntensity # for summarizing several datasets
```


