---
title: "Using PTRMSR for automatizing pre-processing"
output:
  pdf_document: default
  toc: false
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


\tableofcontents

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This vignette shows how to use PTRMSR package to read .txt files, calculate AUC and return statistical analyses. 

## General principle
In this package, the functions beginning by ptrv use a single dataset as input. The functions beginning by ptrvList use names of several files to be used and a metaData file containing information about the different files. 
The objective of this paper is to use files produced by ptrViewer (or any other preprocessing of PTR-MS device returning files with ion, intensity and time as columns) then return curves and relevant statistics on the evaluations. 
Several biases can be corrected:  (i) the breathing of the subject by calculating statistics by breathing cycle (use the parameter correction="cycle"), (ii) blank period during the evaluation that can be removed from the intensities (use blankCorrection = TRUE and noisePeriod=...). 
Some statistics are returned.

## Loading the library
```{r,eval=FALSE}
library(PTRMSR)
```

## Loading metadata and list of files to be processed
All the data files (.txt from ptrViewer) and the metadata file (.csv) should be in a single repository 
```{r, eval=TRUE}
repo="./../inst/extdata"
listFiles=list.files(repo,pattern="*.txt")
metaData=read.table(paste0(repo,"/metaData.csv"),sep=";",header=T)
head(metaData)
```



## AUC calculations of several PTR-Viewer files with noise and breathing correction
This function allows statistics to be calculated for each file after breathing correction
```{r,eval=FALSE,warning=FALSE,message=FALSE,include=FALSE}
setwd(repo)
res_auc=ptrvListIntensityByTime(listFiles=listFiles,
                                metaData=metaData,ions=c("m31.0183", "m31.0259", "m33.0324"))
head(res_auc$listRes)
```

The results can be saved into csv files
```{r,eval=FALSE}
write.table(file="auc.csv",sep=";",res_auc$listRes,row.names=F)
```

The stat option can be used to select the statistic.
```{r,eval=FALSE}
res_tmax=ptrvListIntensityByTime(listFiles=listFiles,metaData=metaData,ions=c("m31.0183", "m31.0259",
                                                                              "m33.0324"),stat="tmax")
```

The breathing cycles can be observed with this command.
```{r,eval=FALSE}
grid.arrange(grobs=res_tmax$cycleLimits[1:4])
```

## Documentations on the functions are available in the R documentation
```{r,eval=FALSE}
?detectCycle
?ptrvIntensityByTime # for only one dataset, 
?ptrvIntensity # for summarizing the time of one dataset
?ptrvListIntensityByTime # for only one dataset, 
?ptrvListIntensity # for summarizing several datasets
```


