---
title: "VignettePTRViewer for sta"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\tableofcontents 

# Installation and requirements
This vignette shows how to use PTRMSR package to detect the start of peaks. 
First, the package should be installed. 

```{r, eval=FALSE,echo=TRUE,warning=FALSE,message=FALSE}
library(devtools)
install_github("www.github.com/ChemoSens/PTRMSR")
```

If required, the commands to load MSnbase are : 
```{r, eval=FALSE,echo=TRUE,warning=FALSE,message=FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("MSnbase")
```

This vignette uses PTRMSR, but also other packages that should be installed and loaded.
```{r, eval=FALSE,echo=TRUE,warning=FALSE,message=FALSE}
install.packages(c("pheatmap","openxlsx","ggplot2","signal","gridExtra")
```

```{r, eval=TRUE,echo=FALSE,warning=FALSE,message=FALSE}
library(PTRMSR)
library(pheatmap)
library(ggplot2)
library(signal)
library(gridExtra)
library(MSnbase)
library(openxlsx)
``` 

Some further functions are available for these vignette, but not in the package
(see .Rmd).
```{r,echo=FALSE, eval=TRUE}
showSurimposedData=function(subFiles,smooth=TRUE,nPoints=31)
{
  noise=rep(NA,length(subFiles));names(noise)=subFiles
dfgg=NULL
for(file in subFiles)
{
  ptrv=read.table(file,sep="\t",header=T)
  longDf=ptrv[,c("RelTime","m131.09549..Ethyl.2methylbutyrate...Conc.")]
  longDf[,"ion"]="m131.09549..Ethyl.2methylbutyrate...Conc."
  longDf[,"file"]=file
  colnames(longDf)=c("time","intensity","ion","file")
 
  if(smooth)
  {
      sgolay0=sgolayfilt(x=longDf[,"intensity"], p = 3,n=nPoints,m=0)
      longDf[,"intensity"]=sgolay0
      longDf[longDf[,"intensity"]<0,"intensity"]=0
  }
      
  #library(plotly); ggplotly(p)
  sp1 <- new("Spectrum1",
           intensity = longDf[,"intensity"],
           mz = longDf[,"time"],
           centroided = FALSE)
  dfgg=rbind(dfgg,longDf)
  noise[file]=estimateNoise(sp1, method = "MAD")[1,"intensity"]
}
noiseMax=3*max(noise)
p=ggplot(dfgg,aes(x=time,y=intensity,color=file))+geom_line()+
  geom_hline(yintercept=noiseMax)+theme_bw()
 return(list(noise=noise,dfgg=dfgg,p=p))
}


totalResultsFun=function(files,method,ion,nPoints=31,startPeaks,statOfNoise="blperiod",
                         smooth=TRUE,proportionOfMax=0.0032,peakChoice="firstAmongHigh",
                         detectionThreshold=3,multiplyNoiseBy=2.7,thresholdOut=0.3,
                         firstAmongHighThreshold=10,
  stat3=read.xlsx(xlsxFile="C:/Users/capeltier/Desktop/DataAnalysis/Christopher PTR/Caroline_Peltier/Statistique_3.xlsx"))
{
    finalMatrix=matrix(NA,length(files),length(startPeaks));
    rownames(finalMatrix)=files;colnames(finalMatrix)=paste0("pic",1:length(startPeaks))
    finalMatrix2=finalMatrix3=finalMatrix
    cv5=rep(NA,length(files));names(cv5)=files
   # pb = txtProgressBar(min = 1, max = length(files), initial = 1) 
    nbOut0=nbPicProb=nbOut=nbNa0=nbNaManual=nbPicVraimentProb=cv5
    column=paste0("diff_",method)
    results=data.frame()
    df=NULL
    i=1
    for(file in files)
    {
      # print(file)
        if(!is.null(stat3)){
            manual=data.frame(X1=as.numeric(stat3[stat3[,"Fichier"]==substr(file,1,23),"Latence"])/1000)
        }
        if(is.null(stat3)){manual=NULL}
        ptrFile_i=read.table(file,sep="\t",header=T)
        startPeaks=startPeaks
        resds=detectStartsInFile(ptrFile_i,ion=ion,startPeaks=startPeaks,
                                 endPeaks=NULL,methods=method,
                                 nPoints=nPoints,multiplyNoiseBy=multiplyNoiseBy,
                                 statOfNoise=statOfNoise,smooth=smooth,
                                 proportionOfMax=proportionOfMax,peakChoice=peakChoice,
                                 detectionThreshold=detectionThreshold,
                                 firstAmongHighThreshold=firstAmongHighThreshold,manual=manual)
     #   setTxtProgressBar(pb,i)
      
        print(paste0(i,"/",length(files)))
        i=i+1
        df_file=as.data.frame(resds$dfres)
        df_file[,"file"]=file
        df=rbind(df,df_file)
       # print("ok")
    }
    
    return(df)
}

``` 
# First observations for settings "smooth' 'firstAmongPeakThreshold' and 'detectionThreshold'

## Selecting detectionThreshold and smoothing (to be done simultaneously: smoothing reduces intensities)
```{r, eval=TRUE,warning=FALSE}
setwd("C:/Users/capeltier/Desktop/DataAnalysis/Christopher PTR/Caroline_Peltier/A1")
files=list.files()

subFiles=files[c(1,100,150)]

res61=showSurimposedData(subFiles,smooth=TRUE,nPoints=61)
res31=showSurimposedData(subFiles,smooth=TRUE,nPoints=31)
res0=showSurimposedData(subFiles,smooth=FALSE)

grid.arrange(res61$p+ggtitle("61 points of smooth")+
               xlim(102,106),res31$p +ggtitle("31 points of smooth")+
               xlim(102,106),res0$p+ggtitle("no smooth")+xlim(102,106))
```
The showSurimposedData also return the estimated noise for each file

```{r, eval=TRUE,warning=FALSE}
res31$noise
mean(res31$noise)*1.6
```
## Selecting the firstAmongHigh threshold
This is a priori threshold: default to 10. If 0, the first peak is taken. If 100, only the max peak is taken. 

# ptrvDetectStart: visualising a start detection for a given peak
One start can be detected with the ptrvDetectStart function with several options. This function takes a data.frame whose colnames are 'time','intensity' and 'ion' as input. 

```{r, eval=TRUE}
data(longDf)
summary(longDf)
```

```{r, eval=TRUE}
ion="m131.09549..Ethyl.2methylbutyrate...Conc."
smooth=TRUE
nPoints=31
detectionThreshold=3*0.75
```
Most of the parameters of the functions are described in the help of the function. 
```{r, eval=TRUE}
res1=ptrvDetectStart(res=longDf,starts=ion,startPeriod=c(156,176),method="higherThanNoise",
                       multiplyNoiseBy=3, noisePeriod=c(156,176),statOfNoise="blperiod",
                       peakChoice="firstAmongHigh",nPoints=nPoints,smooth=smooth,
                        firstAmongHighThreshold=50,detectionThreshold=detectionThreshold)
```

The start of the peak can be found in tx object. 
```{r, eval=TRUE}
res1$tx
```

The results can be plotted. 
```{r, eval=TRUE,warning=F,fig.height=2,fig.width=5}
res1$gg+xlim(156,168)+ylim(0,150)
```

The plot can be interactive by running

```{r, eval=FALSE}
library(plotly)
ggplotly(res1$gg+xlim(156,176))+ylim(0,150)
```
A diagnosis of the peak (as 'too low') is also returned
```{r, eval=TRUE}
res1$diagnosis
```

If the firstAmongHighThreshold option is modified (from 50% to 1%), the chosen peak is not the same: 
```{r, eval=TRUE,warning=F,fig.height=2,fig.width=5}
res2=ptrvDetectStart(res=longDf,starts=ion,startPeriod=c(156,176),method="higherThanNoise",
                     multiplyNoiseBy=3,noisePeriod=c(156,176),statOfNoise="blperiod",
                     peakChoice="firstAmongHigh",nPoints=nPoints,smooth=smooth,
                     firstAmongHighThreshold=1,detectionThreshold=detectionThreshold)
res2$gg+xlim(156,168)+ylim(0,150)
```

This specific case (the first peak is not the highest) is displayed in diagnosis. 

```{r, eval=TRUE}
res2$diagnosis
```

The potential peaks can be also displayed
```{r, eval=TRUE}
res2$potentialPeaks
```

The calculated intensity to be reached to be the beginning of the peak is stored in intensityThreshold.
```{r, eval=TRUE}
res2$intensityThreshold
```


# detectStartsInFile: Detecting starts with a given sequency of stimulations
A function detectStartsInFile detects several starts given a vector of stimulations (called startPeaks)
```{r, eval=TRUE}
data(startPeaks)
head(startPeaks)
data(ptrvFile_i)
ptrvFile_i[1:4,1:5]
```

This function takes the same parameters as ptrvDetectStart and run it on several intervals given by startPeaks. The starting periods are constituted of the consecutive startPeaks. It also allows the methods to be compared.
```{r, eval=TRUE,warning=FALSE}
res=detectStartsInFile(ptrFile_i=ptrvFile_i,ion=ion,startPeaks=startPeaks,
        methods=c("startPeakProportion","higherThanNoise"),
        nPoints=31,smooth=TRUE,
        multiplyNoiseBy=2.576 ,statOfNoise="blperiod",
        proportionOfMax=0.005,
        peakChoice="firstAmongHigh",detectionThreshold=detectionThreshold,
        firstAmongHighThreshold=10)

```

Results of detectStartsInFile contains different elements stored in a list
```{r, eval=TRUE}
names(res)
```
They can be plotted
```{r, eval=TRUE,fig.height=2.5}
res$gg
```

```{r, eval=TRUE,warning=FALSE}
res=detectStartsInFile(ptrFile_i=ptrvFile_i,ion=ion,startPeaks=startPeaks,
        methods=c("startPeakProportion","higherThanNoise"),
        nPoints=31,smooth=TRUE,
        multiplyNoiseBy=2.576 ,statOfNoise="blperiod",
        proportionOfMax=0.005,
        peakChoice="firstAmongHigh",detectionThreshold=detectionThreshold,
        firstAmongHighThreshold=10)

```
The results can be summarized. 
```{r, eval=TRUE}
summary(res$dfres)
```

# Getting all problematic peaks for several files
```{r, eval=FALSE}
setwd("C:/Users/capeltier/Desktop/DataAnalysis/Christopher PTR/Caroline_Peltier/A1")
filesToTest=list.files()
restest=totalResultsFun(files=filesToTest,nPoints=31,statOfNoise="blperiod"
                              ,smooth=TRUE,ion=ion,proportionOfMax=0.005,
                              method="higherThanNoise",peakChoice="firstAmongHigh",
                              startPeaks=startPeaks,detectionThreshold=1.5,
                              multiplyNoiseBy=1.6,stat3=NULL,
                        firstAmongHighThreshold=10)
summary(restest)
summary(factor(restest$diagnosis))
```


```{r, eval=TRUE}
restest[restest$diagnosis%in%c("maxPeakTooLow"),c("pic","file")]
```

# Comparing startPeakProportion and higherThanNoise
## Getting a pattern for higherThanNoise method
```{r, eval=FALSE,warning=FALSE}
setwd("C:/Users/capeltier/Desktop/DataAnalysis/Christopher PTR/Caroline_Peltier/A1")
filesToTest=files
mnb=seq(1.63,1.74,length.out=20);resnoise=list()
for(i in 1:length(mnb))
{
resnoise[[i]]=totalResultsFun(filesToTest,nPoints=31,statOfNoise="blperiod"
                              ,smooth=TRUE,ion=ion,proportionOfMax=0.005,
                              method="higherThanNoise",peakChoice="firstAmongHigh",
                              startPeaks=startPeaks,detectionThreshold=detectionThreshold,
                              multiplyNoiseBy=mnb[i],stat3=NULL,
                        firstAmongHighThreshold=10)
}
```
## Getting a pattern for peakProportion method

```{r, eval=FALSE,warning=FALSE}
respeak=list()
pom=seq(0.0064,0.0072,length.out=20)
for(i in 1:length(mnb))
{
respeak[[i]]=totalResultsFun(files=filesToTest,nPoints=31,statOfNoise="blperiod",
              smooth=TRUE,ion=ion,startPeaks=startPeaks,proportionOfMax=pom[i],
              method="startPeakProportion",peakChoice="firstAmongHigh",
              detectionThreshold=detectionThreshold,multiplyNoiseBy=2.7,
                        firstAmongHighThreshold=10,stat3=NULL)
}
```

## Obtaining the median difference between the two methods
```{r, eval=FALSE,warning=FALSE}
difference=matrix(NA,length(mnb),length(pom))
colnames(difference)=round(100*pom,digits=2)
rownames(difference)=round(mnb,digits=2)
for(i in 1:length(mnb))
{
  for(j in 1:length(pom))
  {
      difference[i,j]=median(respeak[[j]][,"resp"]-resnoise[[i]][,"resp_higherThanNoise"],
                             na.rm=T)
  }
}
```
## Plotting the corresponding image
```{r, eval=FALSE,warning=FALSE}
library(pheatmap)
pheatmap(difference,cluster_rows=F,cluster_cols=F,
         main="Differences between \n startPeakProportion and higherThanNoise")

pheatmap(abs(difference),cluster_rows=F,cluster_cols=F,
         main="Absolute differences between \n startPeakProportion and higherThanNoise")
```
## The minimal difference median : for which values ?
```{r, eval=FALSE,warning=FALSE}
indexes=which(abs(difference)==min(abs(difference)),arr.ind=TRUE)
min(abs(difference))
mnb[indexes[1]]
pom[indexes[2]]

```

## Number of problematic peaks (and their relative files)
```{r, eval=FALSE,warning=FALSE}
nbNaMnb=rep(NA,length(mnb));names(nbNaMnb)=rownames(difference)
nbNaPom=rep(NA,length(pom));names(nbNaPom)=colnames(difference)
for(i in 1:length(mnb))
{
  nbNaMnb[i]=sum(is.infinite(resnoise[[i]][,"resp_higherThanNoise"]))
}
for(j in 1:length(pom))
{
  nbNaPom[j]=sum(is.infinite(respeak[[j]][,"resp"]))
}
```

# Comparison to manual results with detectStartsInFile (and result)

Results can be compared to results obtained manually
```{r, eval=TRUE,warning=FALSE}
data(manual)
head(manual)
res_manual=detectStartsInFile(ptrFile_i=ptrvFile_i,ion=ion,startPeaks=startPeaks,
        methods=c("startPeakProportion","higherThanNoise"),
        nPoints=31,smooth=TRUE,
        multiplyNoiseBy=2.576 ,statOfNoise="blperiod",
        proportionOfMax=0.005,
        peakChoice="firstAmongHigh",detectionThreshold=detectionThreshold,
        firstAmongHighThreshold=10,
        manual=manual)
```
The results can be summarized. 
```{r, eval=TRUE}
summary(res_manual$dfres)
```
To observe the median of the absolute value of differences between automatic and manual results for "startPeakProportion" and "higherThanNoise", type:
```{r, eval=TRUE}
median(abs(res_manual$dfres$diff_startPeakProportion),na.rm=T)
median(abs(res_manual$dfres$diff_higherThanNoise),na.rm=T)
```


## Optimizing the parameters
### Finding the proportion of intensity maximizing the similarities with manual
```{r, eval=FALSE}
setwd("C:/Users/capeltier/Desktop/DataAnalysis/Christopher PTR/Caroline_Peltier/A1")
proportionOfMaxTests=seq(0,0.02,length.out=10)
med_peak=rep(NA,length(proportionOfMaxTests));names(med)=round(proportionOfMaxTests,digits=3)
manuals=peakDetected=c()
for(i in 1:length(med))
{
  respeak=totalResultsFun(files[c(1,100,150)],method="startPeakProportion",
              nPoints=31,statOfNoise="blperiod",smooth=TRUE,
              proportionOfMax=proportionOfMaxTests[i],
              peakChoice="firstAmongHigh",detectionThreshold=detectionThreshold,
              multiplyNoiseBy=2.7,  firstAmongHighThreshold=10)
  med_peak[i]=median(abs(respeak$diff_startPeakProportion),na.rm=T)
  peakDetected=c(peakDetected,respeak$resp)
  manuals=c(manuals,manual)
}
plot(proportionOfMaxTests,med,type="l")
plot(med_peak,type="l")
```

### Finding the number "multiplyByNoise" maximizing the similarities with manual
```{r, eval=FALSE}
setwd("C:/Users/capeltier/Desktop/DataAnalysis/Christopher PTR/Caroline_Peltier/A1")
multiplyNoiseByTests=seq(0.1,5,length.out=10)
manuals=noiseDetected=c()
med_noise=rep(NA,length(proportionOfMaxTests));names(med)=round(proportionOfMaxTests,digits=2)
for(i in 1:length(med))
{
  resnoise=totalResultsFun(files[c(1,100,150)],method="higherThanNoise",nPoints=31,
                           statOfNoise="blperiod",smooth=TRUE,
                        multiplyNoiseBy=multiplyNoiseByTests[i],
                          peakChoice="firstAmongHigh",detectionThreshold=detectionThreshold,
                        firstAmongHighThreshold=10)
  med_noise[i]=median(abs(resnoise$diff_higherThanNoise),na.rm=T)
  noiseDetected=c(noiseDetected,resnoise$resp)
  manuals=c(manuals,manual)
}
plot(med_noise,type="l")
```


